\documentclass[11pt,a4paper]{article}

\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage[margin=2.5cm]{geometry}
\usepackage{hyperref}
\usepackage{xcolor}
\usepackage{listings}
\usepackage{enumitem}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{amsmath,amssymb}
\usepackage{fancyhdr}
\usepackage{titlesec}

\definecolor{isls-blue}{HTML}{1a5276}
\definecolor{isls-orange}{HTML}{e67e22}
\definecolor{isls-green}{HTML}{27ae60}
\definecolor{isls-red}{HTML}{c0392b}
\definecolor{isls-gray}{HTML}{7f8c8d}
\definecolor{code-bg}{HTML}{f7f9fb}

\lstdefinestyle{rust}{
  language=C,
  basicstyle=\ttfamily\small,
  keywordstyle=\color{isls-blue}\bfseries,
  commentstyle=\color{isls-gray}\itshape,
  stringstyle=\color{isls-green},
  backgroundcolor=\color{code-bg},
  frame=single,
  rulecolor=\color{isls-gray!30},
  breaklines=true,
  showstringspaces=false,
  tabsize=4,
  morekeywords={fn,let,mut,pub,struct,impl,enum,use,mod,self,Self,
    crate,super,where,trait,for,in,if,else,match,return,async,await,
    Ok,Err,Some,None,Vec,String,HashMap,HashSet,PathBuf,Result,Option,
    f64,usize,bool,u32,u64,Box,dyn},
}
\lstdefinestyle{bash}{
  language=bash,
  basicstyle=\ttfamily\small,
  backgroundcolor=\color{code-bg},
  frame=single,
  rulecolor=\color{isls-gray!30},
  breaklines=true,
  showstringspaces=false,
}

\pagestyle{fancy}
\fancyhf{}
\fancyhead[L]{\small\textcolor{isls-gray}{ISLS M1 Spec --- MERKABA Orbitkern}}
\fancyhead[R]{\small\textcolor{isls-gray}{v2.0.0 --- \today}}
\fancyfoot[C]{\thepage}

\titleformat{\section}
  {\Large\bfseries\color{isls-blue}}{\thesection}{1em}{}
\titleformat{\subsection}
  {\large\bfseries\color{isls-blue!80}}{\thesubsection}{1em}{}
\titleformat{\subsubsection}
  {\normalsize\bfseries\color{isls-blue!60}}{\thesubsubsection}{1em}{}

\hypersetup{
  colorlinks=true,
  linkcolor=isls-blue,
  urlcolor=isls-orange,
  citecolor=isls-green,
}

\begin{document}

\begin{titlepage}
\centering
\vspace*{2.5cm}

{\Huge\bfseries\color{isls-blue} ISLS M1 Specification}\\[0.5cm]
{\LARGE\color{isls-orange} MERKABA Orbitkern}\\[0.3cm]
{\large\color{isls-gray} Backend-agnostischer Ophanim Swarm\\
mit Resonanz-Konsensus f\"ur Code-Generierung}\\[2cm]

{\large
\begin{tabular}{rl}
\textbf{System:}       & ISLS --- Intelligent Semantic Ledger Substrate \\
\textbf{Version:}      & v5.0 (M1) \\
\textbf{Author:}       & Sebastian Klemm \\
\textbf{Date:}         & \today \\
\textbf{Status:}       & Specification \\
\textbf{Depends:}      & D8 (complete), PSP-1.2-SRI-MERKABA v3.0 \\
\textbf{Branch:}       & \texttt{claude/implement-isls-m1-spec} \\
\textbf{Crate:}        & \texttt{isls-merkaba} (new) \\
\end{tabular}
}

\vfill

{\small\color{isls-gray}
Mithras Substructure University\\
Repository: \texttt{https://github.com/LashSesh/graphity}\\
Reference: PSP Schnitt 1.2 --- Structural Resonance Intelligence (Klemm, April 2026)\\
License: MIT
}
\end{titlepage}

\tableofcontents
\newpage


% ══════════════════════════════════════════════════════════════════════════════
\section{Executive Summary}
% ══════════════════════════════════════════════════════════════════════════════

After D8, ISLS calls one oracle once per LLM-node. If the answer is 
bad, S6 Coagula repairs after the fact. This is reactive.

M1 replaces the single call with the \textbf{Ophanim Swarm}: $n$ 
calls to \emph{any} oracle (OpenAI, Ollama, anything implementing 
the \texttt{Oracle} trait), each with a different context projection, 
aggregated through structural resonance consensus, and collapsed 
into a single emission only when coherence exceeds a threshold.

The key architectural decision: \texttt{SwarmOracle} wraps a 
\texttt{Box<dyn Oracle>}. It does not know or care what oracle it 
wraps. The consensus math is identical whether the inner oracle is 
GPT-4o at \$0.60/swarm-run or a local 7B model at \$0.00. 
The intelligence is in the exoskeleton, not the oracle.

\subsection{What M1 Produces}

A new crate \texttt{isls-merkaba} containing:

\begin{enumerate}[nosep]
\item \textbf{SwarmOracle:} wraps any \texttt{Box<dyn Oracle>}, 
      calls it $n$ times with $n$ different projections, applies 
      resonance consensus, returns the best answer. Drop-in 
      replacement --- implements \texttt{Oracle}.
\item \textbf{Chameleon Pipeline:} deterministic prompt projector. 
      One prompt $\to$ $n$ focused variants. Zero LLM cost.
\item \textbf{Ophanim Coherer:} structural code similarity across 
      $n$ candidates. Per-Thronengel resonance product $D_k$.
\item \textbf{Konus Lens:} $D_\text{total}$ from geometric mean 
      of all $D_k$. Soft unanimity.
\item \textbf{Monolith Gate:} Mandorla selection (best candidate) 
      or Swarm-Coagula (retry with evidence).
\end{enumerate}

No dependency on Ollama, OpenAI, or any specific provider. 
\texttt{isls-merkaba} depends only on \texttt{isls-forge-llm} 
(for the \texttt{Oracle} trait) and \texttt{isls-reader} 
(for Rust code parsing).


% ══════════════════════════════════════════════════════════════════════════════
\section{Core Design: Oracle Wrapping}
% ══════════════════════════════════════════════════════════════════════════════

\begin{lstlisting}[style=rust]
/// The Ophanim Swarm Oracle.
///
/// Wraps ANY Oracle implementation. Calls it n times with
/// n different Chameleon projections. Applies resonance
/// consensus. Returns the Mandorla candidate.
///
/// The inner oracle does not know it is inside a swarm.
/// The caller does not know a swarm was used.
pub struct SwarmOracle {
    /// The wrapped oracle (OpenAI, Ollama, Mock, anything).
    inner: Box<dyn Oracle>,
    /// Number of Thronengel per call (default: 4).
    pub swarm_size: usize,
    /// Resonance threshold Theta (default: 0.25).
    pub threshold: f64,
    /// Max Swarm-Coagula rounds (default: 2).
    pub max_coagula: u32,
}

impl SwarmOracle {
    pub fn new(inner: Box<dyn Oracle>, swarm_size: usize) -> Self {
        Self {
            inner,
            swarm_size,
            threshold: 0.25,
            max_coagula: 2,
        }
    }
    
    pub fn with_threshold(mut self, theta: f64) -> Self {
        self.threshold = theta;
        self
    }
}

impl Oracle for SwarmOracle {
    fn call(
        &self,
        prompt: &str,
        max_tokens: u32,
    ) -> Result<String, Box<dyn std::error::Error>> {
        // 1. Chameleon: prompt -> n projections
        let projections = chameleon::project(prompt, self.swarm_size);
        
        // 2. Swarm: call inner oracle n times
        let mut candidates: Vec<String> = Vec::new();
        for pi in &projections {
            match self.inner.call(pi, max_tokens) {
                Ok(resp) => candidates.push(resp),
                Err(e) => {
                    tracing::warn!("Thronengel failed: {}", e);
                    candidates.push(String::new());
                }
            }
        }
        
        // 3. Ophanim: measure coherence
        let dk = ophanim::compute_resonance(&candidates);
        
        // 4. Konus: D_total
        let (dtotal, omega) = konus::compute_dtotal(&dk);
        
        // 5. Monolith: decide
        monolith::select(
            &candidates, &dk, dtotal, omega,
            self.threshold, self.max_coagula,
            self.inner.as_ref(),
            prompt, max_tokens,
        )
    }
}
\end{lstlisting}

\subsection{Composability Example}

\begin{lstlisting}[style=rust]
// Single OpenAI call (current behavior):
let oracle = Box::new(OpenAiOracle::new(key, "gpt-4o"));

// Swarm of 4 OpenAI calls with consensus:
let oracle = Box::new(SwarmOracle::new(
    Box::new(OpenAiOracle::new(key, "gpt-4o")),
    4,
));

// Swarm of 4 Ollama calls with consensus:
let oracle = Box::new(SwarmOracle::new(
    Box::new(OllamaOracle::new("codellama:7b", "http://localhost:11434")),
    4,
));

// Swarm of 2 on weak hardware:
let oracle = Box::new(SwarmOracle::new(
    Box::new(OllamaOracle::new("phi3:mini", "http://localhost:11434")),
    2,
));

// All four pass to StagedClosure identically.
// StagedClosure does not know which variant it has.
\end{lstlisting}


% ══════════════════════════════════════════════════════════════════════════════
\section{The Chameleon Pipeline}
\label{sec:chameleon}
% ══════════════════════════════════════════════════════════════════════════════

\subsection{Four Canonical Lenses}

Each Thronengel receives the full base prompt plus a focus directive. 
The directives are engineering-level instructions, not vague hints:

\begin{longtable}{clp{7cm}}
\toprule
\textbf{$k$} & \textbf{Lens} & \textbf{Directive (appended to prompt)} \\
\midrule
1 & \textbf{Correctness}
  & ``PRIORITY: Type correctness. Every \texttt{use} import MUST 
    resolve. Every function signature MUST match the AVAILABLE 
    IMPORTS section exactly. If an import path is not listed, 
    do NOT use it. Prefer explicit types over inference.'' \\
2 & \textbf{Completeness}
  & ``PRIORITY: Complete implementation. Every CRUD function must 
    have a full body. No \texttt{todo!()}, no \texttt{unimplemented!()}, 
    no stub returns. Every SQL query must include all fields from 
    the model struct.'' \\
3 & \textbf{Robustness}
  & ``PRIORITY: Error handling. Every \texttt{sqlx::query} call 
    must \texttt{.map\_err(|e| AppError::DatabaseError(...))}. 
    Not-found must return \texttt{AppError::NotFound}. Duplicate 
    key must return \texttt{AppError::Conflict}. Never 
    \texttt{.unwrap()} on database results.'' \\
4 & \textbf{Consistency}
  & ``PRIORITY: Naming conventions. Functions: 
    \texttt{get\_\{entity\}}, \texttt{list\_\{entities\}}, 
    \texttt{create\_\{entity\}}, \texttt{update\_\{entity\}}, 
    \texttt{delete\_\{entity\}}. Parameters: 
    \texttt{pool: \&PgPool}, then payload, then id. Match the 
    style of the PROVIDED SYMBOLS exactly.'' \\
\bottomrule
\end{longtable}

These four lenses map to the four most common failure modes 
in ISLS code generation (identified from D1--D7 debug cycles):

\begin{enumerate}[nosep]
\item Wrong imports (Rule 11, ProvidedSymbol mismatch).
\item Incomplete implementations (\texttt{todo!()} stubs).
\item Missing error handling (\texttt{.unwrap()} on queries).
\item Wrong naming (Rule 3 violations).
\end{enumerate}

Each lens attacks one failure mode. The consensus filters out 
answers that fail on any of the four.

\subsection{Implementation}

\begin{lstlisting}[style=rust]
pub fn project(base_prompt: &str, n: usize) -> Vec<String> {
    let lenses = [
        LENS_CORRECTNESS,
        LENS_COMPLETENESS,
        LENS_ROBUSTNESS,
        LENS_CONSISTENCY,
    ];
    
    (0..n).map(|k| {
        let lens = lenses[k % lenses.len()];
        format!(
            "{base}\n\n\
             ## FOCUS DIRECTIVE (Thronengel {idx}/{total})\n\
             {lens}\n\n\
             Produce ONLY the Rust source code. \
             No markdown fences. No explanations.",
            base = base_prompt,
            idx = k + 1,
            total = n,
            lens = lens,
        )
    }).collect()
}
\end{lstlisting}

Deterministic. Same prompt $\to$ same projections. 
For $n = 1$: returns one prompt with Correctness lens 
(graceful degradation to single-call-plus-focus).


% ══════════════════════════════════════════════════════════════════════════════
\section{The Ophanim Coherer}
\label{sec:ophanim}
% ══════════════════════════════════════════════════════════════════════════════

\subsection{Structural Code Similarity}

Given code strings $c_1, \ldots, c_n$, extract from each using 
\texttt{isls\_reader::parse\_string(c\_k, Language::Rust)}:

\begin{itemize}[nosep]
\item $F(c_k)$: function names with arity (\texttt{get\_trade/2}).
\item $S(c_k)$: struct/enum names.
\item $I(c_k)$: \texttt{use} import paths.
\item $L(c_k)$: line count.
\end{itemize}

Pairwise similarity (Barbara weights from \texttt{isls-code-topo}):

\begin{equation}
\text{sim}(a, b) = 0.40 \cdot J(F_a, F_b) + 0.25 \cdot J(S_a, S_b) 
                  + 0.20 \cdot J(I_a, I_b) + 0.15 \cdot \frac{\min(L_a, L_b)}{\max(L_a, L_b)}
\label{eq:sim}
\end{equation}

\subsection{Per-Ophanim Resonance Product}

\begin{equation}
\psi_k = \frac{1}{n-1} \sum_{j \neq k} \text{sim}(c_k, c_j) \quad\text{(agreement)}
\end{equation}
\begin{equation}
\rho_k = \min\!\left(1.0,\;\frac{|F(c_k)| + |S(c_k)|}{5}\right) \quad\text{(content richness, capped)}
\end{equation}
\begin{equation}
\omega_k = \begin{cases} 
1.0 & \text{if } c_k \text{ has } \geq 1 \text{ function AND } \geq 1 \text{ use statement} \\ 
0.5 & \text{if only partial structure} \\ 
0.0 & \text{if empty or no Rust structure found}
\end{cases}
\end{equation}
\begin{equation}
D_k = \psi_k \cdot \rho_k \cdot \omega_k
\end{equation}

If $\psi_k < 0.15$ (extreme outlier): $D_k = 0$.


% ══════════════════════════════════════════════════════════════════════════════
\section{The Konus Lens}
\label{sec:konus}
% ══════════════════════════════════════════════════════════════════════════════

\begin{equation}
D_\text{total} = \left(\prod_{k=1}^{n} \max(D_k, \epsilon)\right)^{1/n} \cdot \Omega
\label{eq:dtotal}
\end{equation}

where $\epsilon = 10^{-6}$ prevents log-domain collapse, and:

\begin{equation}
\Omega = \frac{2}{n(n-1)} \sum_{i < j} \text{sim}(c_i, c_j)
\quad\text{(mean pairwise similarity)}
\end{equation}

Properties (from SRI Paper \S6.3):
\begin{itemize}[nosep]
\item \textbf{Soft unanimity:} If any $D_k \to 0$, 
      $D_\text{total} \to 0$.
\item \textbf{Noise suppression:} Variance scales as $O(1/n)$ 
      with swarm size (Theorem 6.7).
\item \textbf{False positive rate} $\to 0$ as $n \to \infty$ 
      (Corollary C.1).
\end{itemize}

Default threshold $\Theta = 0.20$, tunable via 
\texttt{--swarm-threshold}.


% ══════════════════════════════════════════════════════════════════════════════
\section{The Monolith Gate}
\label{sec:monolith}
% ══════════════════════════════════════════════════════════════════════════════

\subsection{Mandorla Selection}

If $D_\text{total} \geq \Theta$: return the candidate with the 
highest $D_k$. This is the Thronengel that agrees most with the 
group, has the richest content, and parses correctly. The 
\emph{Mandorla} --- the region where all perspectives overlap.

\subsection{Swarm-Coagula}

If $D_\text{total} < \Theta$:

\begin{enumerate}[nosep]
\item Take the candidate with the highest $D_k$ as base.
\item Identify the disagreements: functions present in $\geq$2 
      candidates but absent from the base.
\item Build a Coagula prompt: base code + ``The following elements 
      were expected by peer review but are missing: [list]. 
      Fix the code to include them.''
\item Call \texttt{self.inner.call(coagula\_prompt, max\_tokens)}.
\item Re-evaluate coherence with the new answer included.
\item Max \texttt{self.max\_coagula} rounds (default 2).
\end{enumerate}

If Coagula exhausts: return the best candidate anyway. Graceful 
degradation (Rule 10). The downstream S5 Gate (cargo check) 
catches remaining issues.

\subsection{Logging}

\begin{lstlisting}[style=bash]
[MERKABA] Swarm(n=4, oracle=OpenAI/gpt-4o) for norm_queries.rs
  T1 (Correctness):  D=0.72  fns=5 structs=2 imports=8
  T2 (Completeness): D=0.68  fns=5 structs=2 imports=7
  T3 (Robustness):   D=0.61  fns=4 structs=2 imports=8
  T4 (Consistency):  D=0.70  fns=5 structs=2 imports=8
  Omega=0.82  Dtotal=0.55  Theta=0.20 -> EMIT (Mandorla: T1)
  Tokens: 4x2340=9360  Time: 12.4s
\end{lstlisting}


% ══════════════════════════════════════════════════════════════════════════════
\section{Work Packages}
% ══════════════════════════════════════════════════════════════════════════════

\subsection{W1: Crate isls-merkaba}

\subsubsection{File Structure}

\begin{lstlisting}[style=bash]
crates/isls-merkaba/
  Cargo.toml
  src/
    lib.rs           # pub mod + SwarmOracle re-export
    chameleon.rs     # Chameleon Pipeline (4 lenses)
    ophanim.rs       # Structural similarity + Dk
    konus.rs         # Dtotal, Omega, threshold
    monolith.rs      # Mandorla selection, Swarm-Coagula
    swarm_oracle.rs  # Oracle trait impl, orchestration
\end{lstlisting}

\subsubsection{Dependencies}

\begin{lstlisting}[style=bash]
[dependencies]
isls-forge-llm = { path = "../isls-forge-llm" }  # Oracle trait
isls-reader    = { path = "../isls-reader" }      # parse_string
serde          = { workspace = true }
serde_json     = { workspace = true }
tracing        = { workspace = true }
# NO reqwest — inner Oracle handles all HTTP
\end{lstlisting}

Note: \texttt{isls-merkaba} has \textbf{no HTTP dependency}. 
It never makes network calls. It calls \texttt{self.inner.call()}, 
and the inner oracle handles the transport. This makes 
\texttt{isls-merkaba} a pure consensus engine, testable 
without any server.

\subsubsection{W1 Verification}

\begin{enumerate}[nosep]
\item \texttt{cargo build -p isls-merkaba} --- compiles.
\item Unit test: \texttt{chameleon::project(prompt, 4)} returns 
      4 strings, each containing one of the 4 focus directives.
\item Unit test: \texttt{ophanim::compute\_resonance()} on 4 
      identical code strings returns all $D_k > 0.8$.
\item Unit test: 4 completely different strings $\to$ all 
      $D_k < 0.2$.
\item Unit test: \texttt{konus::compute\_dtotal()} with uniform 
      $D_k = 0.7$ returns $D_\text{total} > 0.4$.
\item Unit test: \texttt{konus::compute\_dtotal()} with one 
      $D_k = 0.0$ returns $D_\text{total} \approx 0$.
\item Unit test: \texttt{monolith::select()} picks highest-$D_k$.
\item Unit test: SwarmOracle with MockOracle($n=4$) returns 
      the mock's response (consensus trivial when identical).
\end{enumerate}


\subsection{W2: CLI Integration}

\subsubsection{Commands}

\begin{lstlisting}[style=bash]
# Swarm with OpenAI (uses API budget):
isls forge-chat -m "Pet shop" --swarm --api-key $KEY

# Swarm with Ollama (free, local):
isls forge-chat -m "Pet shop" --swarm --ollama

# Customize swarm:
isls forge-chat -m "Pet shop" --swarm --swarm-size 6 \
    --swarm-threshold 0.30 --api-key $KEY

# Single call (default, unchanged behavior):
isls forge-chat -m "Pet shop" --api-key $KEY
\end{lstlisting}

Oracle selection (extended):
\begin{enumerate}[nosep]
\item \texttt{--mock-oracle} $\to$ MockOracle
\item \texttt{--swarm --api-key} $\to$ SwarmOracle(OpenAiOracle)
\item \texttt{--swarm --ollama} $\to$ SwarmOracle(OllamaOracle)
\item \texttt{--swarm} alone $\to$ SwarmOracle(MockOracle) --- useful 
      for testing consensus math without cost
\item \texttt{--ollama} $\to$ OllamaOracle (single call)
\item \texttt{--api-key} $\to$ OpenAiOracle (single call)
\item None $\to$ MockOracle
\end{enumerate}

\subsubsection{Modified Files}

\begin{longtable}{lp{10cm}}
\toprule
\textbf{File} & \textbf{Change} \\
\midrule
\texttt{isls-cli/src/main.rs}
  & \texttt{--swarm}, \texttt{--swarm-size}, 
    \texttt{--swarm-threshold} flags. SwarmOracle construction. \\
\texttt{isls-cli/Cargo.toml}
  & Add \texttt{isls-merkaba} dependency. \\
\texttt{Cargo.toml} (workspace)
  & Add \texttt{isls-merkaba} to members. \\
\bottomrule
\end{longtable}


\subsection{W3: Metrics + Verification}

\subsubsection{Extended Metrics}

\begin{lstlisting}[style=rust]
pub struct GenerationMetrics {
    // ... existing fields from D7 ...
    pub oracle_type: String,     // "mock"|"openai"|"ollama"|"swarm"
    pub swarm_size: Option<usize>,
    pub swarm_inner: Option<String>,  // "openai"|"ollama"|"mock"
    pub swarm_dtotal_avg: Option<f64>,
    pub swarm_emit_rate: Option<f64>, // % of nodes where Monolith fired
    pub swarm_coagula_rounds: Option<u32>,
}
\end{lstlisting}

\subsubsection{Comparison Report}

\begin{lstlisting}[style=bash]
isls metrics --compare

ISLS Generation Metrics — Oracle Comparison
---
                   Single     Swarm(4)
Generations:       12         5
Compile success:   83%        100%
Avg coagula (S6):  0.42       0.00
Avg Dtotal:        --         0.58
Emit rate:         --         100%
Avg tokens:        12,400     49,600
Avg duration:      180s       380s
\end{lstlisting}

The hypothesis: Swarm uses $4\times$ tokens and $2\times$ time but 
achieves $\sim$100\% compile success with zero S6 Coagula, because 
the resonance consensus pre-filters errors before they reach 
\texttt{cargo check}.


% ══════════════════════════════════════════════════════════════════════════════
\section{Cost Model}
% ══════════════════════════════════════════════════════════════════════════════

\begin{longtable}{lrrr}
\toprule
\textbf{Mode} & \textbf{Calls/Run} & \textbf{Cost/Run} & \textbf{Runs/\$20} \\
\midrule
Single OpenAI    & 7    & \$0.15  & 133 \\
Swarm(4) OpenAI  & 28   & \$0.60  & 33 \\
Swarm(4) Ollama  & 28   & \$0.00  & $\infty$ \\
Swarm(2) OpenAI  & 14   & \$0.30  & 66 \\
\bottomrule
\end{longtable}

With \$20 OpenAI budget: 33 full Swarm runs. Enough to verify 
the hypothesis. When budget is tight: \texttt{--swarm-size 2} 
halves the cost while still providing two independent perspectives 
(Correctness + Completeness --- the two most impactful lenses).


% ══════════════════════════════════════════════════════════════════════════════
\section{Constraints for Claude Code}
% ══════════════════════════════════════════════════════════════════════════════

\begin{enumerate}
\item \texttt{isls-merkaba} has \textbf{zero HTTP dependencies}. 
      No reqwest. No tokio. No network code. It calls 
      \texttt{self.inner.call()} and that is all.
\item \texttt{SwarmOracle} MUST implement \texttt{Oracle} from 
      \texttt{isls-forge-llm}. No trait changes.
\item \texttt{SwarmOracle::new(inner, swarm\_size)} takes 
      \texttt{Box<dyn Oracle>}. The inner oracle is opaque.
\item Chameleon projections MUST be deterministic.
\item Code parsing via \texttt{isls\_reader::parse\_string()}.
      Do not reimplement Rust parsing.
\item Graceful degradation everywhere: failed inner call $\to$ 
      empty candidate (not panic). Exhausted Coagula $\to$ return 
      best (not error).
\item Do NOT modify isls-forge-llm, isls-norms, isls-reader, 
      isls-code-topo, isls-gateway, or any structural generators.
\item All existing tests MUST pass.
\item Branch: \texttt{claude/implement-isls-m1-spec}.
\item Commit format: \texttt{M1/W\{n\}: <description>}.
\end{enumerate}


% ══════════════════════════════════════════════════════════════════════════════
\section{Post-M1 Roadmap}
% ══════════════════════════════════════════════════════════════════════════════

\begin{description}[nosep]
\item[M2 --- ScorpioSync:] Phase-tagged event loop. Thronengel 
  activate in phase windows. Temporal coordination.
\item[M3 --- Gabriel Cells:] $(\psi, \rho, \omega)$ extraction 
  at input level. Norm-matching becomes resonance-based.
\item[M4 --- MEF Ledger:] Hash-chained evidence replacing flat 
  \texttt{norms.json}. Every Monolith event, every norm 
  promotion recorded with cryptographic integrity.
\item[M5 --- Adaptive $\Theta$:] Threshold learns from history. 
  High-SGB domains get stricter consensus. New domains get 
  more exploration.
\end{description}

\vfill
\begin{center}
\color{isls-gray}\rule{0.5\textwidth}{0.4pt}\\[0.5em]
{\small End of M1 Specification v2.0.\\[0.3em]
The oracle is replaceable.\\
The exoskeleton is the intelligence.\\
--- PSP Schnitt 1.2, \S1}
\end{center}

\end{document}
