\documentclass[11pt,a4paper]{article}

\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage[margin=2.5cm]{geometry}
\usepackage{hyperref}
\usepackage{xcolor}
\usepackage{listings}
\usepackage{enumitem}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{amsmath,amssymb}
\usepackage{fancyhdr}
\usepackage{titlesec}

\definecolor{isls-blue}{HTML}{1a5276}
\definecolor{isls-orange}{HTML}{e67e22}
\definecolor{isls-green}{HTML}{27ae60}
\definecolor{isls-red}{HTML}{c0392b}
\definecolor{isls-gray}{HTML}{7f8c8d}
\definecolor{code-bg}{HTML}{f7f9fb}

\lstdefinestyle{rust}{
  language=C,
  basicstyle=\ttfamily\small,
  keywordstyle=\color{isls-blue}\bfseries,
  commentstyle=\color{isls-gray}\itshape,
  stringstyle=\color{isls-green},
  backgroundcolor=\color{code-bg},
  frame=single,
  rulecolor=\color{isls-gray!30},
  breaklines=true,
  showstringspaces=false,
  tabsize=4,
  morekeywords={fn,let,mut,pub,struct,impl,enum,use,mod,self,Self,
    crate,super,where,trait,for,in,if,else,match,return,async,await,
    Ok,Err,Some,None,Vec,String,HashMap,HashSet,PathBuf,Result,Option,
    f64,usize,bool,u32,u64},
}
\lstdefinestyle{bash}{
  language=bash,
  basicstyle=\ttfamily\small,
  backgroundcolor=\color{code-bg},
  frame=single,
  rulecolor=\color{isls-gray!30},
  breaklines=true,
  showstringspaces=false,
}

\pagestyle{fancy}
\fancyhf{}
\fancyhead[L]{\small\textcolor{isls-gray}{ISLS M1 Spec --- MERKABA Orbitkern}}
\fancyhead[R]{\small\textcolor{isls-gray}{v1.0.0 --- \today}}
\fancyfoot[C]{\thepage}

\titleformat{\section}
  {\Large\bfseries\color{isls-blue}}{\thesection}{1em}{}
\titleformat{\subsection}
  {\large\bfseries\color{isls-blue!80}}{\thesubsection}{1em}{}
\titleformat{\subsubsection}
  {\normalsize\bfseries\color{isls-blue!60}}{\thesubsubsection}{1em}{}

\hypersetup{
  colorlinks=true,
  linkcolor=isls-blue,
  urlcolor=isls-orange,
  citecolor=isls-green,
}

\begin{document}

\begin{titlepage}
\centering
\vspace*{2.5cm}

{\Huge\bfseries\color{isls-blue} ISLS M1 Specification}\\[0.5cm]
{\LARGE\color{isls-orange} MERKABA Orbitkern}\\[0.3cm]
{\large\color{isls-gray} Ophanim Swarm, Chameleon Pipeline,\\
Resonanz-Konsensus f\"ur Code-Generierung}\\[2cm]

{\large
\begin{tabular}{rl}
\textbf{System:}       & ISLS --- Intelligent Semantic Ledger Substrate \\
\textbf{Version:}      & v5.0 (M1) \\
\textbf{Author:}       & Sebastian Klemm \\
\textbf{Date:}         & \today \\
\textbf{Status:}       & Specification \\
\textbf{Depends:}      & D8 (Ollama Oracle), PSP-1.2-SRI-MERKABA v3.0 \\
\textbf{Branch:}       & \texttt{claude/implement-isls-m1-spec} \\
\textbf{Crate:}        & \texttt{isls-merkaba} (new) \\
\end{tabular}
}

\vfill

{\small\color{isls-gray}
Mithras Substructure University\\
Repository: \texttt{https://github.com/LashSesh/graphity}\\
Reference: PSP Schnitt 1.2 --- Structural Resonance Intelligence (Klemm, April 2026)\\
License: MIT
}
\end{titlepage}

\tableofcontents
\newpage


% ══════════════════════════════════════════════════════════════════════════════
\section{Executive Summary}
% ══════════════════════════════════════════════════════════════════════════════

After D8, ISLS calls one oracle once per LLM-node: one prompt, one 
answer, one file. If the answer is bad, S6 Coagula tries to fix it 
after the fact. This is reactive repair, not proactive quality.

M1 replaces the single oracle call with the \textbf{Ophanim Swarm}: 
$n$ parallel Ollama instances, each receiving a different context 
projection (Chameleon Pipeline), aggregated through resonance 
consensus (Konus), and collapsed into a single emission only when 
system-wide coherence exceeds a gated threshold (Monolith).

The oracle is not the intelligence. The exoskeleton is.

\subsection{What M1 Produces}

A new crate \texttt{isls-merkaba} containing:

\begin{enumerate}[nosep]
\item \textbf{SwarmOracle:} implements the \texttt{Oracle} trait.
      Drop-in replacement for \texttt{OllamaOracle}. 
      Externally identical: \texttt{call(prompt, max\_tokens) 
      $\to$ String}. Internally: $n$ parallel calls, resonance 
      consensus, Mandorla selection.
\item \textbf{Chameleon Pipeline:} deterministic prompt projector.
      Takes one base prompt and produces $n$ variants, each 
      focusing on a different architectural concern.
\item \textbf{Ophanim Coherer:} measures structural similarity 
      across $n$ code candidates. Computes per-Ophanim resonance 
      products $D_k$.
\item \textbf{Konus Lens:} computes $D_\text{total}$ from all 
      $D_k$. Determines if consensus is sufficient.
\item \textbf{Monolith Gate:} selects the Mandorla candidate 
      (closest to centroid) or triggers Swarm-level Coagula.
\end{enumerate}

\subsection{The Key Constraint: Same Trait, No Pipeline Change}

\texttt{SwarmOracle} implements \texttt{Oracle}. The 
\texttt{StagedClosure} calls \texttt{oracle.call(prompt, 
max\_tokens)} exactly as before. It does not know whether a 
single Ollama or 4 Thronengel answered. The MERKABA exoskeleton 
is invisible to the pipeline.

This means: zero changes to \texttt{hdag.rs}, 
\texttt{staged\_closure.rs}, \texttt{structural.rs}, 
\texttt{static\_files.rs}, or any other forge module. M1 is 
purely additive.


% ══════════════════════════════════════════════════════════════════════════════
\section{Architecture}
% ══════════════════════════════════════════════════════════════════════════════

\subsection{Data Flow: Single Oracle Call (pre-M1)}

\begin{lstlisting}[style=bash]
StagedClosure S4 Solve (per LLM node):
  prompt = build_hdag_prompt(node, provided, type_ctx, plan)
  response = oracle.call(prompt, 4096)
  write_file(node.path, response)
\end{lstlisting}

\subsection{Data Flow: Ophanim Swarm (M1)}

\begin{lstlisting}[style=bash]
SwarmOracle::call(prompt, max_tokens):
  |
  +-- Chameleon: prompt -> [pi_1, pi_2, ..., pi_n]
  |     (n context projections)
  |
  +-- Swarm Dispatch: parallel Ollama calls
  |     response_1 = ollama.call(pi_1, max_tokens)
  |     response_2 = ollama.call(pi_2, max_tokens)
  |     ...
  |     response_n = ollama.call(pi_n, max_tokens)
  |
  +-- Ophanim Coherer: measure structural similarity
  |     D_1 = coherence(response_1, all_responses)
  |     D_2 = coherence(response_2, all_responses)
  |     ...
  |     D_n = coherence(response_n, all_responses)
  |
  +-- Konus: D_total = geometric_mean(D_1..D_n) * Omega
  |
  +-- Monolith: D_total >= Theta?
  |     YES -> select Mandorla candidate -> return
  |     NO  -> Coagula (feed back, retry, max 2 rounds)
  |            -> if exhausted: return best candidate
  |
  +-- return: String (single best response)
\end{lstlisting}

The caller sees: \texttt{call(prompt) $\to$ String}. Identical 
interface. Radically different internal process.


% ══════════════════════════════════════════════════════════════════════════════
\section{The Chameleon Pipeline}
\label{sec:chameleon}
% ══════════════════════════════════════════════════════════════════════════════

\subsection{Purpose}

Each Thronengel sees the same problem from a different angle. This 
is not random variation --- it is structured context projection. 
The projections are deterministic functions of the base prompt.

\subsection{Projection Lenses}

For code generation, four canonical lenses (from SRI Paper \S6.2):

\begin{longtable}{clp{7cm}}
\toprule
\textbf{$k$} & \textbf{Lens} & \textbf{What the Thronengel Focuses On} \\
\midrule
1 & \textbf{Correctness}
  & ``Focus on type correctness. Every import must resolve. 
    Every function signature must match the provided symbols 
    exactly. Prefer safe, simple code over clever code.'' \\
2 & \textbf{Completeness}
  & ``Ensure every CRUD operation is fully implemented. No 
    stub functions. No todo comments. Every code path must 
    return a meaningful value or error.'' \\
3 & \textbf{Robustness}
  & ``Focus on error handling. Every database call must have 
    error mapping. Every user input must be validated. Every 
    edge case (empty list, not found, duplicate) must be 
    handled.'' \\
4 & \textbf{Consistency}
  & ``Follow the naming conventions exactly. Function names: 
    get\_\{entity\}, list\_\{entities\}, create/update/delete. 
    Use the exact types from the provided symbols. Match the 
    style of the other files in the project.'' \\
\bottomrule
\end{longtable}

\subsection{Prompt Construction}

\begin{lstlisting}[style=rust]
pub fn chameleon(
    base_prompt: &str,
    n: usize,
) -> Vec<String> {
    let lenses = &[
        LENS_CORRECTNESS,
        LENS_COMPLETENESS,
        LENS_ROBUSTNESS,
        LENS_CONSISTENCY,
    ];
    
    (0..n).map(|k| {
        let lens = lenses[k % lenses.len()];
        format!(
            "{}\n\n## FOCUS DIRECTIVE (Thronengel {})\n{}\n\
             Produce ONLY the Rust source code. No markdown.",
            base_prompt, k + 1, lens
        )
    }).collect()
}
\end{lstlisting}

For $n = 4$: each gets a unique lens. For $n > 4$: lenses cycle 
(Thronengel 5 gets lens 1 again, providing a second perspective 
on correctness). For $n = 1$: falls back to the base prompt 
with lens 1 (Correctness) --- degrading gracefully to single-call.

\subsection{Determinism}

The Chameleon Pipeline is \textbf{fully deterministic}: same prompt 
$\to$ same projections. No randomness. No LLM involved. This 
preserves PSP replay-stability (SI4).


% ══════════════════════════════════════════════════════════════════════════════
\section{The Ophanim Coherer}
\label{sec:ophanim}
% ══════════════════════════════════════════════════════════════════════════════

\subsection{Purpose}

Measure how much $n$ independently generated code files agree 
structurally. Not string equality --- structural agreement.

\subsection{Structural Similarity for Rust Code}

Given two code strings $a$ and $b$, extract:

\begin{itemize}[nosep]
\item $F(x)$: set of function signatures (\texttt{pub fn name(params) -> ReturnType}).
\item $S(x)$: set of struct names.
\item $I(x)$: set of \texttt{use} imports.
\item $L(x)$: line count.
\end{itemize}

Pairwise similarity:

\begin{equation}
\text{sim}(a, b) = 0.40 \cdot J(F(a), F(b)) + 0.25 \cdot J(S(a), S(b)) 
                  + 0.20 \cdot J(I(a), I(b)) + 0.15 \cdot \text{size\_sim}(L(a), L(b))
\label{eq:sim}
\end{equation}

where $J$ is Jaccard similarity and $\text{size\_sim}(a, b) = 
\min(a,b) / \max(a,b)$.

These are the same weights as \texttt{topology\_similarity()} in 
\texttt{isls-code-topo} (``hardened weights from Barbara, 95\% 
cross-language recognition'').

\subsection{Per-Ophanim Resonance Product}

For Thronengel $k$ with code $c_k$:

\begin{equation}
\psi_k = \text{mean}_{j \neq k} \text{sim}(c_k, c_j) \quad\text{(agreement with others)}
\end{equation}
\begin{equation}
\rho_k = \frac{|F(c_k)| + |S(c_k)|}{|F(c_k)| + |S(c_k)| + \epsilon} \quad\text{(content density; $\epsilon$ prevents div/0 on empty)}
\end{equation}
\begin{equation}
\omega_k = \begin{cases} 1.0 & \text{if } c_k \text{ parses as valid Rust (no syntax errors)} \\ 0.5 & \text{if some syntax issues} \\ 0.0 & \text{if empty or unparseable} \end{cases}
\end{equation}
\begin{equation}
D_k = \psi_k \cdot \rho_k \cdot \omega_k
\label{eq:dk}
\end{equation}

If $\psi_k < \eta_\text{thr}$ (default 0.3): $D_k = 0$ (this 
Thronengel is incoherent with the group).

\subsection{Implementation Note: Reuse isls-reader}

The extraction of $F(x)$, $S(x)$, $I(x)$ reuses 
\texttt{isls\_reader::parse\_string(code, Language::Rust)}. The 
code-topo similarity weights are imported from 
\texttt{isls-code-topo}. M1 does not reinvent parsing.


% ══════════════════════════════════════════════════════════════════════════════
\section{The Konus Lens}
\label{sec:konus}
% ══════════════════════════════════════════════════════════════════════════════

\subsection{Total Resonance}

\begin{equation}
D_\text{total} = \left(\prod_{k=1}^{n} D_k\right)^{1/n} \cdot \Omega
\label{eq:dtotal}
\end{equation}

where $\Omega$ is the global convergence field:

\begin{equation}
\Omega = \text{mean}_{(i,j), i < j} \text{sim}(c_i, c_j) \quad\text{(average pairwise similarity)}
\end{equation}

The geometric mean (Proposition 6.6 from SRI Paper) implements 
\textbf{soft unanimity}: if any $D_k = 0$, $D_\text{total} = 0$. 
A single strong answer cannot override incoherent consensus.

\subsection{Threshold}

Default $\Theta = 0.25$ (calibrated lower than the SRI Paper's 0.45 
because code generation has more structural variance than natural 
language). Adjustable via \texttt{--swarm-threshold}.


% ══════════════════════════════════════════════════════════════════════════════
\section{The Monolith Gate}
\label{sec:monolith}
% ══════════════════════════════════════════════════════════════════════════════

\subsection{Decision}

If $D_\text{total} \geq \Theta$: select the \textbf{Mandorla 
candidate} --- the response with the highest individual $D_k$. 
This is the Thronengel that most agrees with the group AND has 
the richest content AND parses correctly.

If $D_\text{total} < \Theta$: \textbf{Swarm-Coagula}. Feed the 
best candidate (highest $D_k$) plus the violation evidence 
(what the others disagreed on) back to a single oracle call. 
Max 2 Swarm-Coagula rounds. If still below $\Theta$: return the 
best candidate anyway (graceful degradation, Rule 10).

\subsection{Logging}

Every Swarm cycle logs to tracing:

\begin{lstlisting}[style=bash]
[MERKABA] Swarm(n=4) for database/norm_queries.rs
  Thronengel 1 (Correctness): D=0.72, 5 fns, 2 structs
  Thronengel 2 (Completeness): D=0.68, 5 fns, 2 structs
  Thronengel 3 (Robustness):   D=0.61, 4 fns, 2 structs
  Thronengel 4 (Consistency):  D=0.70, 5 fns, 2 structs
  Omega=0.82, Dtotal=0.55, Theta=0.25 -> EMIT (Mandorla: T1)
\end{lstlisting}


% ══════════════════════════════════════════════════════════════════════════════
\section{Work Packages}
% ══════════════════════════════════════════════════════════════════════════════

% ──────────────────────────────────────────────────────────────────────────────
\subsection{W1: Crate isls-merkaba}
\label{sec:w1}
% ──────────────────────────────────────────────────────────────────────────────

\subsubsection{Goal}

New crate with the four MERKABA modules. Pure Rust, no async 
(Ollama calls are blocking via \texttt{reqwest::blocking}, 
consistent with existing \texttt{OpenAiOracle}).

\subsubsection{File Structure}

\begin{lstlisting}[style=bash]
crates/isls-merkaba/
  Cargo.toml
  src/
    lib.rs           # pub mod + SwarmOracle re-export
    chameleon.rs     # Chameleon Pipeline
    ophanim.rs       # Ophanim Coherer + Dk computation
    konus.rs         # Dtotal, Omega, threshold
    monolith.rs      # Mandorla selection, Swarm-Coagula
    swarm_oracle.rs  # Oracle trait impl, orchestration
\end{lstlisting}

\subsubsection{Dependencies}

\begin{lstlisting}[style=toml]
[dependencies]
isls-forge-llm = { path = "../isls-forge-llm" }  # Oracle trait
isls-reader    = { path = "../isls-reader" }      # Rust parsing
serde          = { workspace = true }
serde_json     = { workspace = true }
tracing        = { workspace = true }
reqwest        = { workspace = true }
\end{lstlisting}

\subsubsection{SwarmOracle Public API}

\begin{lstlisting}[style=rust]
pub struct SwarmOracle {
    pub swarm_size: usize,       // n (default 4)
    pub threshold: f64,          // Theta (default 0.25)
    pub max_coagula: u32,        // max swarm-coagula rounds (default 2)
    pub ollama_url: String,      // Ollama base URL
    pub model: String,           // Ollama model name
}

impl SwarmOracle {
    pub fn new(
        swarm_size: usize,
        model: &str,
        ollama_url: &str,
    ) -> Self;
}

impl Oracle for SwarmOracle {
    fn call(
        &self,
        prompt: &str,
        max_tokens: u32,
    ) -> Result<String, Box<dyn std::error::Error>>;
}
\end{lstlisting}

\subsubsection{W1 Verification}

\begin{enumerate}[nosep]
\item \texttt{cargo build -p isls-merkaba} --- compiles.
\item Unit test: \texttt{chameleon(prompt, 4)} produces 4 
      distinct prompts with 4 different focus directives.
\item Unit test: \texttt{ophanim\_coherence()} on 4 identical 
      strings returns $D_k \approx 1.0$ for all.
\item Unit test: \texttt{ophanim\_coherence()} on 4 completely 
      different strings returns $D_k \approx 0.0$.
\item Unit test: \texttt{konus\_dtotal()} with all $D_k = 0.7$ 
      returns $D_\text{total} > 0$.
\item Unit test: Mandorla selects highest-$D_k$ candidate.
\end{enumerate}


% ──────────────────────────────────────────────────────────────────────────────
\subsection{W2: CLI Integration}
\label{sec:w2}
% ──────────────────────────────────────────────────────────────────────────────

\subsubsection{Goal}

Add \texttt{--swarm} flag to all forge commands. When active, 
use \texttt{SwarmOracle} instead of \texttt{OllamaOracle}.

\subsubsection{Commands}

\begin{lstlisting}[style=bash]
isls forge-chat -m "Pet shop app" --swarm --model codellama:7b
isls forge-chat -m "Pet shop app" --swarm --swarm-size 6
isls forge-chat -m "Pet shop app" --swarm --swarm-threshold 0.35
isls forge-self --swarm
\end{lstlisting}

Oracle selection priority (extended from D8):
\begin{enumerate}[nosep]
\item \texttt{--mock-oracle} $\to$ MockOracle.
\item \texttt{--swarm} $\to$ SwarmOracle (requires Ollama).
\item \texttt{--ollama} $\to$ OllamaOracle (single-call).
\item \texttt{--api-key} / \texttt{OPENAI\_API\_KEY} $\to$ OpenAiOracle.
\item None $\to$ MockOracle.
\end{enumerate}

\subsubsection{Modified Files}

\begin{longtable}{lp{10cm}}
\toprule
\textbf{File} & \textbf{Change} \\
\midrule
\texttt{isls-cli/src/main.rs}
  & Add \texttt{--swarm}, \texttt{--swarm-size}, 
    \texttt{--swarm-threshold} to forge commands. \\
\texttt{isls-cli/Cargo.toml}
  & Add \texttt{isls-merkaba} dependency. \\
\texttt{Cargo.toml} (workspace)
  & Add \texttt{isls-merkaba} to members. \\
\bottomrule
\end{longtable}


% ──────────────────────────────────────────────────────────────────────────────
\subsection{W3: Metrics + Verification}
\label{sec:w3}
% ──────────────────────────────────────────────────────────────────────────────

\subsubsection{Goal}

Log Swarm metrics into \texttt{GenerationMetrics} (D7 W4) and 
verify that Swarm generation produces higher-quality output 
than single-call.

\subsubsection{New Metric Fields}

\begin{lstlisting}[style=rust]
pub struct GenerationMetrics {
    // ... existing fields ...
    pub oracle_type: String,    // "mock" | "openai" | "ollama" | "swarm"
    pub swarm_size: Option<usize>,
    pub swarm_dtotal_avg: Option<f64>,
    pub swarm_coagula_rounds: Option<u32>,
}
\end{lstlisting}

\subsubsection{Comparison}

\begin{lstlisting}[style=bash]
isls metrics --compare

ISLS Generation Metrics — Oracle Comparison
---
                  Mock    OpenAI   Ollama   Swarm(4)
Generations:      12      8        3        5
Compile success:  100%    83%      67%      100%
Avg coagula:      0.0     0.42     0.67     0.00
Avg Dtotal:       --      --       --       0.61
Avg tokens:       0       12400    8200     32800
Avg duration:     0.3s    180s     240s     480s
\end{lstlisting}

The expected trade-off: Swarm uses $4\times$ more tokens and 
$2\text{--}3\times$ more time, but achieves higher compile 
success rate and zero Coagula cycles (because consensus 
pre-filters bad code before it reaches \texttt{cargo check}).


% ══════════════════════════════════════════════════════════════════════════════
\section{Resource Planning: VPS}
% ══════════════════════════════════════════════════════════════════════════════

For a 4-Ophanim Swarm on a VPS with GPU:

\begin{longtable}{llr}
\toprule
\textbf{Provider} & \textbf{Config} & \textbf{Cost/Month} \\
\midrule
Hetzner CCX33 & 8 vCPU, 32 GB, no GPU & \EUR{30} \\
Hetzner GEX44 & GPU (L40S), 32 GB & \EUR{90} \\
Vast.ai spot & RTX 3090, 24 GB & \textasciitilde\EUR{20--40} \\
RunPod & RTX 3090, 24 GB & \textasciitilde\EUR{30--50} \\
\bottomrule
\end{longtable}

With a 7B model on RTX 3090 (24 GB VRAM): 4 concurrent 
inferences $\times$ 5s each = 20s per Swarm cycle. A 50-node 
HDAG with 7 LLM-nodes: 7 $\times$ 20s = 140s Swarm time + 
structural time. Comparable to current OpenAI latency.

Without GPU (CPU-only, 8 vCPU): 4 sequential inferences 
$\times$ 30--60s = 120--240s per Swarm cycle. Slower but 
functional. The Mandorla selection still works --- it just 
takes longer.

Local on Samsung Galaxy Book4 (8 GB RAM): $n = 2$ 
(sequential, 3B model). Minimal Swarm, still better than 
$n = 1$ because the two lenses (Correctness + Completeness) 
catch different failure modes.


% ══════════════════════════════════════════════════════════════════════════════
\section{Constraints for Claude Code}
% ══════════════════════════════════════════════════════════════════════════════

\begin{enumerate}
\item \texttt{isls-merkaba} is a \textbf{new crate}. Zero changes 
      to existing crates except CLI integration (main.rs) and 
      workspace Cargo.toml.
\item \texttt{SwarmOracle} MUST implement the existing 
      \texttt{Oracle} trait from \texttt{isls-forge-llm}. No 
      trait modification.
\item The Chameleon Pipeline MUST be deterministic. Same prompt 
      $\to$ same projections. No randomness.
\item Ollama HTTP calls use \texttt{reqwest::blocking}. No tokio 
      in isls-merkaba.
\item Graceful degradation: if Swarm-Coagula exhausts, return 
      best candidate (do not fail).
\item \texttt{isls-reader} for Rust parsing, \texttt{isls-code-topo}
      weights for similarity. Import, do not reimplement.
\item All existing tests MUST pass. Pipeline bat MUST still 
      produce 6/6 PASS.
\item Branch: \texttt{claude/implement-isls-m1-spec}.
\item Commit format: \texttt{M1/W\{n\}: <description>}.
\end{enumerate}


% ══════════════════════════════════════════════════════════════════════════════
\section{Dependency Graph}
% ══════════════════════════════════════════════════════════════════════════════

\begin{verbatim}
W1 (isls-merkaba Crate)
  |
  v
W2 (CLI Integration)
  |
  v
W3 (Metrics + Verification)
\end{verbatim}


% ══════════════════════════════════════════════════════════════════════════════
\section{Acceptance Criteria}
% ══════════════════════════════════════════════════════════════════════════════

\begin{longtable}{clp{9cm}}
\toprule
\textbf{\#} & \textbf{Pass} & \textbf{Criterion} \\
\midrule
1 & & \texttt{cargo build --workspace} compiles (incl. isls-merkaba). \\
2 & & \texttt{cargo test --workspace} passes. \\
3 & & \texttt{cargo test -p isls-merkaba} --- Chameleon, Ophanim, 
      Konus, Monolith unit tests pass. \\
4 & & \texttt{isls forge-chat --swarm --mock-oracle} --- runs 
      SwarmOracle with MockOracle internals (no Ollama needed). \\
5 & & With Ollama running: \texttt{isls forge-chat -m "Pet shop" 
      --swarm} generates + compiles. \\
6 & & MERKABA log lines visible in output. \\
7 & & Pipeline bat: 6/6 PASS (backward compat, default path 
      does not use Swarm). \\
8 & & \texttt{isls metrics} shows \texttt{oracle\_type: "swarm"} 
      entries. \\
\bottomrule
\end{longtable}


% ══════════════════════════════════════════════════════════════════════════════
\section{Post-M1: The Exoskeleton Grows}
% ══════════════════════════════════════════════════════════════════════════════

\begin{description}[nosep]
\item[M2 --- ScorpioSync Clock:] Phase-tagged message loop. 
  Thronengel activate in phase order. Enables temporal 
  coordination beyond simple parallelism.
\item[M3 --- Gabriel Cells:] $(\psi, \rho, \omega)$ signature 
  extraction at input level. Norm-matching becomes resonance-based 
  instead of keyword-based.
\item[M4 --- Hash-Chained Ledger:] Replace flat \texttt{norms.json} 
  with MEF evidence chain. Every Monolith event, every norm 
  promotion, every generation is hash-linked and replay-verifiable.
\item[M5 --- Adaptive $\Theta$:] The threshold learns from 
  generation history. Domains with high SGB get higher $\Theta$ 
  (stricter consensus). New domains get lower $\Theta$ 
  (more explorative).
\end{description}

The Donnerkeil built the pipeline. The MERKABA series builds 
the nervous system.

\vfill
\begin{center}
\color{isls-gray}\rule{0.5\textwidth}{0.4pt}\\[0.5em]
{\small End of M1 Specification.\\[0.3em]
A weak oracle inside a strong exoskeleton\\
outperforms a strong oracle inside no exoskeleton.\\
--- PSP Schnitt 1.2, \S1}
\end{center}

\end{document}
